---
title: "Land-use data harmonization"
author: "Vincent Pellissier"
date: "8 November 2018"
header-includes:
   - \usepackage{float}
output:
  pdf_document: default
  html_document: default
classoption: table
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
options(kableExtra.latex.load_packages = FALSE)
options(warn = 0)
library(readxl)
library(tidyverse)
library(magrittr)
library(knitr)
library(kableExtra)
library(float)

path_grassplot <- './'
```
This report explains the steps taken to correct the land-use data.
The masterfile used is the version 1.8, as found on November 8th 2018.
The sheet in Grassplot 1.8.xlsx containing the LU information is the sheet 'datasets' and was read and saved as a .rds file to ease the process (faster loading)

```{r, echo = T}
df <- readRDS(file.path(path_grassplot, 'Grassplot 1.8_Data.rds'))
df <- df %>%
        mutate_at(.vars = c(95:104), funs(ifelse(. %in% c('NA', '[NA]'), NA, .)))
```

The columns 98 (mowing frequency) and 100 (grazing intensity) are renamed:
```{r}
names(df)[c(98,100)] <- c('mowing_frequency', 'grazing_intensity')
```

**PRELIMINARY NOTE: there duplicated ID in the column `Grassplot ID of plot`.
The following ID are duplicated `r unique(df[which(duplicated(df[,3])),3])%>%pull()`**  


## Correction of the land use column
The file lookup_table_LU.xlsx contains two sheets:

  * land_use_detail: lookup table between the column 96 and other column. This needs to be discussed further before I finalize the table.
  * global_land_use: lookuptable matching the values in column 95 with a new classification. This new classification is a set of binary columns
  (allowing for mixed land-use) plus one declarative column. Each binary column correspond to a management practice (grazed,	mown,	burnt,	fertilized, abandoned,	natural_grassland), the declarative column stores additional information that cannot be put in one of the 6 columns (*i.e* trampled) and
  that will be used afterward. **NOTE** LAS, P, L were noted as land-use for the dataset PL_C and N was noted as land-use for the dataset RU_J. I could not find what these stand for. I assumed that these plots don't have a land-use but the info is still in the column 'other'.

```{r}
lut <- read_excel(file.path(path_grassplot, "lookup_table_LU.xlsx" ),
                 sheet = 'global_land_use')
```

The global_land_use lookup table is used to harmonize the column 95 with the new binary columns:
```{r}
df <- df %>%
    left_join(lut, by = c("Land use (5 standard categories: mown, grazed, abandoned, natural grassland, NA)" = "old_classification"))
```

## Correction of the intensity/frequency columns
```{r, echo = F}
grazing_values <- na.omit(unique(df$grazing_intensity))
mowing_values <- na.omit(unique(df$mowing_frequency))
numeric_grazing_values <- na.omit(unique(as.numeric(df$grazing_intensity)))
numeric_mowing_values <- na.omit(unique(as.numeric(df$mowing_frequency)))

non_num_mowing <- setdiff(mowing_values, numeric_mowing_values)
non_num_grazing <- setdiff(grazing_values, numeric_grazing_values)

```

These two columns (mowing_frequency and grazing_intensity) are presently coded as text columns, because they contains odd values.  
The column mowing_frequency contains `r length(non_num_mowing)` non-numeric values `r if(length(non_num_mowing) != 0) {paste0(': ', paste(non_num_mowing, collapse = ','))}`.  
The column grazing_intensity contains `r length(non_num_grazing)` non-numeric values `r if(length(non_num_grazing) != 0) {paste0(': ', paste(non_num_grazing, collapse = ', '))}`.  

### Mowing frequency
The mowing_frequency does not have any odd values.  
The values in the columns are left unchanged but the column is converted to numerical.

```{r}
df$mowing_frequency <- as.numeric(df$mowing_frequency)
```

### Grazing intensity
The grazing_intensity contains a mix of verbal description and numerical values.
The verbal descriptors `r non_num_grazing` are replaced by numerical values and the column is converted to numerical:

  * `overgrazing` and `high` are replaced by 1
  * `middle` is replaced by 0.5
  * `low` is replaced by 0.1

```{r}
df[df$grazing_intensity %in% c('overgrazing', 'high'),]%<>%
  mutate(grazing_intensity = 1)

df[df$grazing_intensity %in% c('middle'),]%<>%
  mutate(grazing_intensity = 0.5)

df[df$grazing_intensity %in% c('low'),]%<>%
  mutate(grazing_intensity = 0.1)

df$grazing_intensity <- as.numeric(df$grazing_intensity)
```


## Matching new binary columns and intensity or duration columns (yet to be automatized)
In this step, we identify datasets with discrepancies between the new binary columns and the matching intensity column to:

  * Manually check in the original datasets or publications
  * Correct the discrepancies.

**NOTE** in the following, I assumed that if the intensity/frequency/duration column was non empty, but the binary was, the intensity/frequency/duration column takes precedence (*i.e* if the frequency/intensity/duration, the binary column HAS to be == 1)
    
### Mowing and mowing intensity
#### *Datasets containing plots for which mowing intensity > 0 & mown != 1 (here and after, we refer to the new binary columns)*
```{r, echo = F}
nb_plots_mown <- sum(df$mown, na.rm = T)

mowing_0_intensity_1 <- df %>% 
          filter(`mown` != 1 & `mowing_frequency` != 0 )


ds_mown <- mowing_0_intensity_1%>%
          distinct(`Dataset ID`)%>%
          pull()

plots_mown <- mowing_0_intensity_1%>%
          distinct(`GrassPlot ID of plot`)%>%
          pull()
```

  * The datasets `r paste(ds_mown, collapse = ', ')` contain `r length(plots_mown)` plots having a mowing intensity but classified as mown == 0. These plots have been manually checked the 8/11/2018 (masterfile v1.8) and contains not errors or additional information. They are thus reclassified as mown == 1.  
  **NOTE** If necessary, we can discuss a frequency threshold the classify a plot as mown or not.
  
```{r}
df[df$mowing_frequency != 0 & !is.na(df$mowing_frequency) & df$mown == 0,] %<>% 
  mutate(mown = 1)
```

**Post reclassification check:** number of plots with a discrepancy - number of plots reclassified (should be 0) = `r length(plots_mown) - (sum(df$mown, na.rm = T) - nb_plots_mown)`


### Grazing and Grazing intensity
```{r, echo = F}
nb_plots_grazed <- sum(df$grazed, na.rm = T)

grazing_0_intensity_1 <- df %>% 
          filter(`grazed` != 1 & `grazing_intensity` != 0 )


ds_grazed <- grazing_0_intensity_1%>%
          distinct(`Dataset ID`)%>%
          pull()

plots_grazed <- grazing_0_intensity_1%>%
          distinct(`GrassPlot ID of plot`)%>%
          pull()
```

#### *Datasets containing plots for which grazing intensity > 0 & grazed != 1:*
  * The datasets `r paste(ds_grazed, collapse = ', ')` contain `r length(plots_grazed)` plots having a mowing intensity but classified as grazed == 0. These plots have been manually checked the 8/11/2018 (masterfile v1.8) and contains not errors or additional information. They are thus reclassified as grazed == 1.  

```{r}
df[df$grazing_intensity != 0 & !is.na(df$grazing_intensity) & df$grazed == 0,] %<>% 
  mutate(grazed = 1)
```

**Post reclassification check:** number of plots with a discrepancy - number of plots reclassified (should be 0) = `r length(plots_grazed) - (sum(df$grazed, na.rm = T) - nb_plots_grazed)` **Manual note: the difference found here is due to duplicates**


## Matching new and old binary columns
In this step, we identify and correct discrepancies between the newly created and corrected binary column, and the former ones.
Here, we consider that mown == 1 or grazed == 1 is always correct, since it is based on the above correction (based either on the broad land-use or on the mowing / grazing intensity).
Thus, we only identify (and potentialy correct) plots for which the new binary column is 0 or NA and the new one is not 0 or NA.

### Mowing
```{r, echo = F}
df%>%
  group_by(`mown`, `Mowing (1/0)`)%>%
  summarise(n=n())%>%
  spread(`mown`, n)%>%
  kable(caption = 'Contingency table of columns mown and Mowing (1/0)')%>%
  kable_styling(latex_options = "HOLD_position")
```

#### Datasets with discrepancies between mown and Mowing (1/0) (automatic report):
```{r, echo = F}
unique_binary <- df %>%
    filter(`mown` == 0 | is.na(mown))%>%
    distinct(`Mowing (1/0)`)%>%
    pull()

unique_binary <- unique_binary[!is.na(unique_binary) & unique_binary != 0]

for(i in unique_binary){
    nb_plots_0 <- df %>%
        filter(`mown` == 0 & `Mowing (1/0)` == i)%>%
        distinct(`GrassPlot ID of plot`)%>% count()
    
    names_datasets_0 <- df %>%
        filter(`mown` == 0 & `Mowing (1/0)` == i)%>%
        distinct(`Dataset ID`)%>% pull()
    
    nb_plots_NA <- df %>%
        filter(is.na(`mown`) & `Mowing (1/0)` == i)%>%
        distinct(`GrassPlot ID of plot`)%>% count()
    
    names_datasets_NA <- df %>%
        filter(is.na(`mown`) & `Mowing (1/0)` == i)%>%
        distinct(`Dataset ID`)%>% pull()
    
    if(nb_plots_0 != 0)
        print(paste("The dataset(s)", paste(names_datasets_0, collapse = ", "), "contain(s)", 
                    nb_plots_0, "plots with mown == 0 and Mowing (1/0) =", i))
    if(nb_plots_NA != 0)
        print(paste("The dataset(s)", paste(names_datasets_NA, collapse = ", "), "contain(s)",
                    nb_plots_NA, "plots with mown = NA and Mowing (1/0) =", i))
}
```

##### Manual verification of discrepancies (has to be updated manually if necessary for each version of the masterfile)
###### *Datasets containing plots with mown = NA & Mowing (1/0) == ? (new vs former column):*
The following plots in PL_A have mown = NA and Mowing (1/0) == '?'
```{r, echo = F}
df %>% 
  filter(is.na(`mown`) & `Mowing (1/0)` == '?')%>%
    distinct(`GrassPlot ID of plot`)%>%
    pull()
```
No additional information could be found, so the plots are left with mown = NA

#### Check
```{r, echo = F}
df%>%
  group_by(`mown`, `Mowing (1/0)`)%>%
  summarise(n=n())%>%
  spread(`mown`, n)%>%
  kable(caption = 'Contingency table of columns mown and Mowing (1/0) after reclassification (check)')%>%
  kable_styling(latex_options = "HOLD_position")
```

### Grazing
```{r, echo = F}
df%>%
  group_by(`grazed`, `Grazing (1/0)`)%>%
  summarise(n=n())%>%
  spread(`grazed`, n)%>%
  kable(caption = 'Contingency table of columns grazed and Grazing (1/0)')%>%
  kable_styling(latex_options = "HOLD_position")
```

#### Datasets with discrepancies between grazed and Grazing (1/0) (automatic report):
```{r, echo = F}
unique_binary <- df %>%
    filter(`grazed` == 0 | is.na(grazed))%>%
    distinct(`Grazing (1/0)`)%>%
    pull()

unique_binary <- unique_binary[!is.na(unique_binary) & unique_binary != 0]

for(i in unique_binary){
    nb_plots_0 <- df %>%
        filter(`grazed` == 0 & `Grazing (1/0)` == i)%>%
        distinct(`GrassPlot ID of plot`)%>% count()
    
    names_datasets_0 <- df %>%
        filter(`grazed` == 0 & `Grazing (1/0)` == i)%>%
        distinct(`Dataset ID`)%>% pull()
    
    nb_plots_NA <- df %>%
        filter(is.na(`grazed`) & `Grazing (1/0)` == i)%>%
        distinct(`GrassPlot ID of plot`)%>% count()
    
    names_datasets_NA <- df %>%
        filter(is.na(`grazed`) & `Grazing (1/0)` == i)%>%
        distinct(`Dataset ID`)%>% pull()
    
    if(nb_plots_0 != 0)
        print(paste("The dataset(s)", paste(names_datasets_0, collapse = ", "), "contain(s)", 
                    nb_plots_0, "plots with grazed == 0 and Grazing (1/0) =", i))
    if(nb_plots_NA != 0)
        print(paste("The dataset(s)", paste(names_datasets_NA, collapse = ", "), "contain(s)",
                    nb_plots_NA, "plots with grazed = NA and Grazing (1/0) =", i))
}
```

##### Manual verification of discrepancies (has to be updated manually if necessary for each version of the masterfile)
###### *Datasets plots with grazed = NA & Grazing (1/0) == '?':*
The following plots in PL_A have grazed = NA and Grazing (1/0) == '?'
```{r echo = F}
df %>%
    filter(is.na(`grazed`) & `Grazing (1/0)` == '?')%>%
    distinct(`GrassPlot ID of plot`)%>% 
    pull()
```
No additional information could be found, so the plots are left as grazed = NA

###### *Datasets plots with grazed == 0 & Grazing (1/0) == 'probably':*
The following plots in PL_A have grazed == 0 and Grazing (1/0) == ''probably'
```{r echo = F}
df %>%
    filter(`grazed` == 0 & `Grazing (1/0)` == 'probably' )%>%
    distinct(`GrassPlot ID of plot`)%>% 
    pull()
```
No additional information could be found, so the plots are left as grazed == 0.


###### *Datasets plots with grazed == 0 & Grazing (1/0) == '1':*
The following plots in PL_A have grazed == 0 and Grazing (1/0) == 1
```{r echo = F}
df %>%
    filter(`grazed` == '0' & `Grazing (1/0)` == '1' )%>%
    distinct(`GrassPlot ID of plot`)%>% 
    pull()
```
  * PL_A contains 6 plots (PL_A_N005_xxxxb) noted as abandoned with 'occasionally grazed or trampled' in the detailed column (column 96). These plots will be classified as grazed in the new binary column.

```{r}
df[df$`grazed` %in% '0' & df$`Grazing (1/0)` %in% '1',]%<>% 
    mutate(grazed = 1)
```

#### Check
```{r, echo = F}
df%>%
  group_by(`grazed`, `Grazing (1/0)`)%>%
  summarise(n=n())%>%
  spread(`grazed`, n)%>%
  kable(caption = 'Contingency table of columns grazed and Grazing (1/0) after reclassification (check')%>%
  kable_styling(latex_options = "HOLD_position")
```



### Burning
```{r, echo = F}
df%>%
  group_by(`burnt`, `Burning (1/0)`)%>%
  summarise(n=n())%>%
  spread(`burnt`, n)%>%
  kable(caption = 'Contingency table of columns burnt and Burning (1/0)')%>%
  kable_styling(latex_options = "HOLD_position")
```

#### Datasets with discrepancies between burnt and Burning (1/0) (automatic report):
```{r, echo = F}
unique_binary <- df %>%
    filter(`burnt` == 0 | is.na(burnt))%>%
    distinct(`Burning (1/0)`)%>%
    pull()

unique_binary <- unique_binary[!is.na(unique_binary) & unique_binary != 0]

for(i in unique_binary){
    nb_plots_0 <- df %>%
        filter(`burnt` == 0 & `Burning (1/0)` == i)%>%
        distinct(`GrassPlot ID of plot`)%>% count()
    
    names_datasets_0 <- df %>%
        filter(`burnt` == 0 & `Burning (1/0)` == i)%>%
        distinct(`Dataset ID`)%>% pull()
    
    nb_plots_NA <- df %>%
        filter(is.na(`burnt`) & `Burning (1/0)` == i)%>%
        distinct(`GrassPlot ID of plot`)%>% count()
    
    names_datasets_NA <- df %>%
        filter(is.na(`burnt`) & `Burning (1/0)` == i)%>%
        distinct(`Dataset ID`)%>% pull()
    
    if(nb_plots_0 != 0)
        print(paste("The dataset(s)", paste(names_datasets_0, collapse = ", "), "contain(s)", 
                    nb_plots_0, "plots with burnt == 0 and Burning (1/0) =", i))
    if(nb_plots_NA != 0)
        print(paste("The dataset(s)", paste(names_datasets_NA, collapse = ", "), "contain(s)",
                    nb_plots_NA, "plots with burnt = NA and Burning (1/0) =", i))
}
```

##### Manual verification of discrepancies (has to be updated manually if necessary for each version of the masterfile)
###### *Datasets with burnt == 0 and Burning (1/0) == 1*
The following plots have burnt = 0 and Burning (1/0) == '1':
```{r echo = F}
df %>%
    filter(burnt == 0 & `Burning (1/0)` == '1' )%>%
    distinct(`GrassPlot ID of plot`)%>% 
    pull()
```

No other information are found. These plots are reclassified with burnt == 1

```{r}
df[df$burnt %in% 0 & df$`Burning (1/0)` %in% '1',]%<>%
    mutate(burnt = 1)
```

###### *Datasets with burnt = NA and Burning (1/0) == 1*
The following plots have burnt = NA and Burning (1/0) == '?':
```{r echo = F}
df %>%
    filter(is.na(burnt) & `Burning (1/0)` == '1')%>%
    distinct(`GrassPlot ID of plot`)%>% 
    pull()
```
No other information are found. These plots are reclassified with burnt == 1

```{r}
df[is.na(df$burnt) & df$`Burning (1/0)` %in% '1',]%<>%
    mutate(burnt = 1)
```


###### *Datasets with burnt = NA and Burning (1/0) == ?*
The following plots have burnt == 0 and Burning (1/0) == '?':
```{r echo = F}
df %>%
    filter(burnt ==0 & `Burning (1/0)` == '?' )%>%
    distinct(`GrassPlot ID of plot`)%>% 
    pull()
```
No additional information could be found, so the plots are left as burnt == 0.

#### Check
```{r, echo = F}
df%>%
  group_by(`burnt`, `Burning (1/0)`)%>%
  summarise(n=n())%>%
  spread(`burnt`, n)%>%
  kable(caption = 'Contingency table of columns burnt and Burning (1/0) after reclassification (check)')%>%
  kable_styling(latex_options = "HOLD_position")
```

### Fertilization
```{r, echo = F}
df%>%
  group_by(`fertilized`, `Fertilized (1/0)`)%>%
  summarise(n=n())%>%
  spread(`fertilized`, n)%>%
  kable(caption = 'Contingency table of columns fertilized and Fertilized (1/0)')%>%
  kable_styling(latex_options = "HOLD_position")
```

This field seems to be a fertilization intensity rather than a binary variable. We should discuss it before making a decision.

## Proposal for a new land-use template
```{r, echo = F}
# Creer un tableau de recap
```

The column 96 provide a lot of information that is so far not fully exploited.
To make use of it, it requires to manually check each of the value in this column, to fill other columns.
After the discussion with Anne and Monika, I propose the following columns:  

### Land-use columns
Already explained above, these are binary columns mown, grazed, burnt, fertilized, abandoned, natural (*i.e. no land-use past or present).
There is one extra column that contains verbal information not found in these five columns.  
**NOTE** This should be discussed, but these columns could also be viewed as management columns.

### Land-cover columns
This columns (binary) could be used to provide a clearer view on the land-use.
After having briefly browsed the column 96, a first proposal is: grassland, moorland, heathland, fallowland, meadow, pasture (not entirely sure about this one).  
**NOTE** Monika was suggesting to find a way to separate primary and secondary grasslands. I'm not sure how this fits in this proposal, so it's open to discussion.

### Time since abandonment
Self explanatory. I'm not quite sure wether this overlap with Ex arable years (years since last ploughing). Maybe that could be combined with an additional column former_land_use, so that will not be restricted to ex-arable plots.

### Grazing animal
Type of animal grazing the land. So far, we have five categories in mind: cow, sheep, goat, horse, other. I am not sure wether grazing by wild animals (which is sometimes present in the land-use details column) should be considered, as it is not necessarily linked with land-use.


Table: Summary of the proposed template (**to be discussed**)

+-------------+------------------+----------------+-------------------+
|Column type  | Column name      |  Variable type | Possible values   |
+=============+==================+================+===================+
| Land use    | - mown           | - binary       | - 0/1             |
|             | - grazed         | - binary       | - 0/1             |
|             | - burnt          | - binary       | - 0/1             |
|             | - fertilized     | - binary       | - 0/1             |
|             | - abandoned      | - binary       | - 0/1             |
|             | - natural        | - binary       | - 0/1             |
|             | - other          | - text         | - undefined       |  
+-------------+------------------+----------------+-------------------+
| Land cover  | - grassland      | - binary       | - 0/1             |
|             | - moorland       | - binary       | - 0/1             |
|             | - heathland      | - binary       | - 0/1             |
|             | - fallowland     | - binary       | - 0/1             |
|             | - meadow         | - binary       | - 0/1             |
|             | - pasture        | - binary       | - 0/1             |
+-------------+------------------+----------------+-------------------+
| Land-use    | - grazing        | - numeric      | - 0 to 1          |
| intensity   | intensity        |                |                   |
|             | - mowing         | - numeric      | - undefined       |
|             | frequency        |                |                   |
|             | - fertilisation  | - numeric      | - undefined       |
|             | intensity        |                |                   |
|             | - time since     | - numeric      | - undefined       |
|             | abandonement     |                |                   |
+-------------+------------------+----------------+-------------------+
| Grazing     | - grazing        | - text         | - cow, sheep,     |
| details     | animal           |                | goat, horse,      |
|             |                  |                | other             |
+-------------+------------------+----------------+-------------------+
