---
title: "Land-use data harmonization"
author: "Vincent Pellissier"
date: "27 November 2018"
header-includes:
   - \usepackage{float}
output:
  pdf_document:
    number_sections: true
  html_document:
    number_sections: true
classoption: table
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
options(kableExtra.latex.load_packages = FALSE)
options(warn = 0)
library(readxl)
library(writexl)
library(tidyverse)
library(magrittr)
library(knitr)
library(kableExtra)
library(float)

path_grassplot <- './'
```
This report explains the steps taken to correct the land-use data.
The masterfile used is the version 1.19, as found on November 27th 2018.
The sheet in Grassplot 1.19.xlsx containing the LU information is the sheet 'datasets' and was read and saved as a .rds file to ease the process (faster loading)

```{r, echo = T}
df <- readRDS(file.path(path_grassplot, 'Grassplot 1.19_Data.rds'))
df <- df %>%
        mutate_at(.vars = c(95:104), funs(ifelse(. %in% c('NA', '[NA]'), NA, .)))
```

The columns 98 (mowing frequency) and 100 (grazing intensity) are renamed as `mowing_frequency` and `grazing intensity`:
```{r}
names(df)[which(names(df) %in% c('Mowing frequency: cuts per year (2=2cut/yr; 1=1cut/yr; 0.5=1 cut/2 yr or 2 yr abandoned; 0.2=1 cut/5 yr or 5 yr abandones; 0=never mown)', 'Grazing intensity (0/1) 1= intensive grazing'))] <- c('mowing_frequency', 'grazing_intensity')
```

The column 104 (Fertilized (1/0)) should be a binary column. Yet, it contains the values `r as.character(na.omit(unique(df[,'Fertilized (1/0)'])))`, meaning that it is an intensity column. The column is thus renamed `fertilization_intensity`
```{r}
names(df)[which(names(df) == 'Fertilized (1/0)')] <- 'fertilization_intensity'
```

**PRELIMINARY NOTE: there are duplicated ID in the column `Grassplot ID of plot`.
The following ID are duplicated `r unique(df[which(duplicated(df[,3])),3])%>%pull()`**  


# Correction of the land use column
The file lookup_table_LU.xlsx contains two sheets:

  * land_use_detail: lookup table between the column 96 and other column. This needs to be discussed further before I finalize the table.
  * global_land_use: lookuptable matching the values in column 95 with a new classification. This new classification is a set of binary columns
  (allowing for mixed land-use) plus one declarative column. Each binary column correspond to a management practice (grazed,	mown,	burnt,	fertilized, abandoned,	natural_grassland), the declarative column stores additional information that cannot be put in one of the 6 columns (*i.e* trampled) and
  that will be used afterward. **NOTE** LAS, P, L were noted as land-use for the dataset PL_C and N was noted as land-use for the dataset RU_J. I could not find what these stand for. I assumed that these plots don't have a land-use but the info is still in the column 'other'.

```{r}
lut <- read_excel(file.path(path_grassplot, "lookup_table_LU.xlsx" ),
                 sheet = 'global_land_use')
```

The global_land_use lookup table is used to harmonize the column 95 with the new binary columns:
```{r}
df <- df %>%
    left_join(lut, 
              by = c("Land use (5 standard categories: mown, grazed, abandoned, natural grassland, NA)" = "old_classification"))
```

# Correction of the intensity/frequency columns
```{r, echo = F}
grazing_values <- na.omit(unique(df$grazing_intensity))
mowing_values <- na.omit(unique(df$mowing_frequency))
fertilization_values <- na.omit(unique(df$fertilization_intensity))
numeric_grazing_values <- na.omit(unique(as.numeric(df$grazing_intensity)))
numeric_mowing_values <- na.omit(unique(as.numeric(df$mowing_frequency)))
numeric_fertilization_values <- na.omit(unique(as.numeric(df$fertilization_intensity)))

non_num_mowing <- setdiff(mowing_values, numeric_mowing_values)
non_num_grazing <- setdiff(grazing_values, numeric_grazing_values)
non_num_fertilization <- setdiff(fertilization_values, numeric_fertilization_values)

nb_mowing <- sum(table(df$mowing_frequency)[names(table(df$mowing_frequency)) != 0])
nb_grazing <- sum(table(df$grazing_intensity)[names(table(df$grazing_intensity)) != 0])
nb_fertilization <- sum(table(df$fertilization_intensity)[names(table(df$fertilization_intensity)) != 0])
```

These three columns (mowing_frequency, grazing_intensity, fertilization_intensity) are presently coded as text columns  
The column mowing_frequency contains `r length(non_num_mowing)` non-numeric values `r if(length(non_num_mowing) != 0) {paste0(': ', paste(non_num_mowing, collapse = ','))}`.  
The column grazing_intensity contains `r length(non_num_grazing)` non-numeric values `r if(length(non_num_grazing) != 0) {paste0(': ', paste(non_num_grazing, collapse = ', '))}`.  
The column fertilization contains `r length(non_num_fertilization)` non-numeric values `r if(length(non_num_fertilization) != 0) {paste0(': ', paste(non_num_fertilization, collapse = ', '))}`. 

## Mowing frequency
The mowing_frequency does have non-numerical values.  
A closer look (manual) at the master files shows that there is some information in the column 96, that can be converted into mowing_frequency.  
The column is converted to numerical, meaning that we lose the information for now, but it will be retrieved later using the column 96.

```{r}
df$mowing_frequency <- as.numeric(df$mowing_frequency)
```

**Post correction check**: original number of plots with frequency - present number of plots with frequency (should be 0) = `r nb_mowing - sum(table(df$mowing_frequency)[names(table(df$mowing_frequency)) != 0])`. **Manual note** The difference here is due to the removal of the plots with mowing_frequency =='x'.  

## Grazing intensity
The grazing_intensity does have non-numerical values.
The verbal descriptors `r non_num_grazing` are replaced by numerical values and the column is converted to numerical:

  * `overgrazing` and `high` are replaced by 1
  * `middle` is replaced by 0.5
  * `low` is replaced by 0.1

```{r}
df[df$grazing_intensity %in% c('overgrazing', 'high'),]%<>%
  mutate(grazing_intensity = 1)

df[df$grazing_intensity %in% c('middle'),]%<>%
  mutate(grazing_intensity = 0.5)

df[df$grazing_intensity %in% c('low'),]%<>%
  mutate(grazing_intensity = 0.1)

df$grazing_intensity <- as.numeric(df$grazing_intensity)
```
**Post correction check**: original number of plots with intensity - present number of plots with intensity (should be 0) = `r nb_grazing - sum(table(df$grazing_intensity)[names(table(df$grazing_intensity)) != 0])`.  

## Fertilization intensity
The fertilization_intensity column does not have non-numerical values.  
The column is converted to numerical.  
```{r}
df$fertilization_intensity <- as.numeric(df$fertilization_intensity)
```
**Post correction check**: original number of plots with intensity - present number of plots with intensity (should be 0) = `r nb_fertilization - sum(table(df$fertilization_intensity)[names(table(df$fertilization_intensity)) != 0])`.  


# Matching new binary columns and frequency/intensity columns.
In this step, we identify datasets with discrepancies between the new binary columns and the matching intensity column to:

  * Manually check in the original datasets or publications
  * Correct the discrepancies.

**NOTE** in the following, I assumed that if the intensity/frequency/duration column was non empty, but the binary was, the intensity/frequency/duration column takes precedence (*i.e* if the frequency/intensity/duration, the binary column **HAS** to be == 1)
    
## Mowing and mowing intensity
```{r, echo = F}
nb_plots_mown <- sum(df$mown, na.rm = T)

mowing_0_intensity_1 <- df %>% 
          filter(`mown` != 1 & `mowing_frequency` != 0 )


ds_mown <- mowing_0_intensity_1%>%
          distinct(`Dataset ID`)%>%
          pull()

plots_mown <- mowing_0_intensity_1%>%
          distinct(`GrassPlot ID of plot`)%>%
          pull()
```

### *Datasets containing plots for which mowing intensity > 0 & mown != 1 (here and after, we refer to the new binary columns)*
  * The datasets `r paste(ds_mown, collapse = ', ')` contain `r length(plots_mown)` plots having a mowing intensity but classified as mown == 0. These plots have been manually checked the 8/11/2018 (masterfile v1.8) and contains not errors or additional information. They are thus reclassified as mown == 1.  
  **NOTE** If necessary, we can discuss a frequency threshold the classify a plot as mown or not.
  
```{r}
df[df$mowing_frequency != 0 & !is.na(df$mowing_frequency) & df$mown == 0,] %<>% 
  mutate(mown = 1)
```

**Post reclassification check:** number of plots with a discrepancy - number of plots reclassified (should be 0) = `r length(plots_mown) - (sum(df$mown, na.rm = T) - nb_plots_mown)`


## Grazing and Grazing intensity
```{r, echo = F}
nb_plots_grazed <- sum(df$grazed, na.rm = T)

grazing_0_intensity_1 <- df %>% 
          filter(`grazed` != 1 & `grazing_intensity` != 0 )


ds_grazed <- grazing_0_intensity_1%>%
          distinct(`Dataset ID`)%>%
          pull()

plots_grazed <- grazing_0_intensity_1%>%
          distinct(`GrassPlot ID of plot`)%>%
          pull()
```

### *Datasets containing plots for which grazing intensity > 0 & grazed != 1:*
  * The datasets `r paste(ds_grazed, collapse = ', ')` contain `r length(plots_grazed)` plots having a mowing intensity but classified as grazed == 0. These plots have been manually checked the 8/11/2018 (masterfile v1.8) and contains not errors or additional information. They are thus reclassified as grazed == 1.  

```{r}
df[df$grazing_intensity != 0 & !is.na(df$grazing_intensity) & df$grazed == 0,] %<>% 
  mutate(grazed = 1)
```

**Post reclassification check:** number of plots with a discrepancy - number of plots reclassified (should be 0) = `r length(plots_grazed) - (sum(df$grazed, na.rm = T) - nb_plots_grazed)`

## Fertilization and fertiliation intensity
```{r, echo = F}
nb_plots_fertilized <- sum(df$fertilized, na.rm = T)

fertilized_0_intensity_1 <- df %>% 
          filter(`fertilized` != 1 & `fertilization_intensity` != 0 )


ds_fertilized <- fertilized_0_intensity_1%>%
          distinct(`Dataset ID`)%>%
          pull()

plots_fertilized <- fertilized_0_intensity_1%>%
          distinct(`GrassPlot ID of plot`)%>%
          pull()
```

### *Datasets containing plots for which fertilization intensity > 0 & fertilized != 1 (here and after, we refer to the new binary columns)*
  * The datasets `r paste(ds_fertilized, collapse = ', ')` contain `r length(plots_fertilized)` plots having a fertilization intensity but classified as fertilized == 0. These plots have **not** been manually verified (to be done). They are thus reclassified as fertilized == 1.  

```{r}
df[df$fertilization_intensity != 0 & !is.na(df$fertilization_intensity) & df$fertilized == 0,] %<>% 
  mutate(fertilized = 1)
```

**Post reclassification check:** number of plots with a discrepancy - number of plots reclassified (should be 0) = `r length(plots_fertilized) - (sum(df$fertilized, na.rm = T) - nb_plots_fertilized)`


# Matching new and old binary columns
In this step, we identify and correct discrepancies between the newly created and corrected binary column, and the former ones.
Here, we consider that mown == 1 or grazed == 1 is always correct, since it is based on the above correction (based either on the broad land-use or on the mowing / grazing intensity).
Thus, we only identify (and potentialy correct) plots for which the old binary column is 0 or NA and the new one is not 0 or NA.

## Mowing
```{r, echo = F}
df%>%
  group_by(`mown`, `Mowing (1/0)`)%>%
  summarise(n=n())%>%
  spread(`mown`, n)%>%
  kable(caption = 'Contingency table of columns mown and Mowing (1/0)')%>%
  kable_styling(latex_options = "HOLD_position")
```

### Datasets with discrepancies between mown and Mowing (1/0) (automatic report):
```{r, echo = F, results = 'asis'}
unique_binary <- df %>%
    filter(`mown` == 0 | is.na(mown))%>%
    distinct(`Mowing (1/0)`)%>%
    pull()

unique_binary <- unique_binary[!is.na(unique_binary) & unique_binary != 0]

for(i in unique_binary){
    nb_plots_0 <- df %>%
        filter(`mown` == 0 & `Mowing (1/0)` == i)%>%
        distinct(`GrassPlot ID of plot`)%>% count()
    
    names_datasets_0 <- df %>%
        filter(`mown` == 0 & `Mowing (1/0)` == i)%>%
        distinct(`Dataset ID`)%>% pull()
    
    nb_plots_NA <- df %>%
        filter(is.na(`mown`) & `Mowing (1/0)` == i)%>%
        distinct(`GrassPlot ID of plot`)%>% count()
    
    names_datasets_NA <- df %>%
        filter(is.na(`mown`) & `Mowing (1/0)` == i)%>%
        distinct(`Dataset ID`)%>% pull()
    
    if(nb_plots_0 != 0){
        cat(paste("\nThe dataset(s)", paste(names_datasets_0, collapse = ", "), "contain(s)", 
                    nb_plots_0, "plots with mown == 0 and Mowing (1/0) =", i))
        cat('\n')
    }
    if(nb_plots_NA != 0){
        cat(paste("\nThe dataset(s)", paste(names_datasets_NA, collapse = ", "), "contain(s)",
                    nb_plots_NA, "plots with mown = NA and Mowing (1/0) =", i))
        cat('\n')
    }
}
```

#### Manual verification of discrepancies (has to be updated manually if necessary for each version of the masterfile)
##### *Datasets containing plots with mown = NA & Mowing (1/0) == ? (new vs former column):*
The following plots in PL_A have mown = NA and Mowing (1/0) == '?'
```{r, echo = F}
df %>% 
  filter(is.na(`mown`) & `Mowing (1/0)` == '?')%>%
    distinct(`GrassPlot ID of plot`)%>%
    pull()
```
No additional information could be found, so the plots are left with mown = NA

#### Check
```{r, echo = F}
df%>%
  group_by(`mown`, `Mowing (1/0)`)%>%
  summarise(n=n())%>%
  spread(`mown`, n)%>%
  kable(caption = 'Contingency table of columns mown and Mowing (1/0) after reclassification (check)')%>%
  kable_styling(latex_options = "HOLD_position")
```

## Grazing
```{r, echo = F}
df%>%
  group_by(`grazed`, `Grazing (1/0)`)%>%
  summarise(n=n())%>%
  spread(`grazed`, n)%>%
  kable(caption = 'Contingency table of columns grazed and Grazing (1/0)')%>%
  kable_styling(latex_options = "HOLD_position")
```

### Datasets with discrepancies between grazed and Grazing (1/0) (automatic report):
```{r, echo = F, results = 'asis'}
unique_binary <- df %>%
    filter(`grazed` == 0 | is.na(grazed))%>%
    distinct(`Grazing (1/0)`)%>%
    pull()

unique_binary <- unique_binary[!is.na(unique_binary) & unique_binary != 0]

for(i in unique_binary){
    nb_plots_0 <- df %>%
        filter(`grazed` == 0 & `Grazing (1/0)` == i)%>%
        distinct(`GrassPlot ID of plot`)%>% count()
    
    names_datasets_0 <- df %>%
        filter(`grazed` == 0 & `Grazing (1/0)` == i)%>%
        distinct(`Dataset ID`)%>% pull()
    
    nb_plots_NA <- df %>%
        filter(is.na(`grazed`) & `Grazing (1/0)` == i)%>%
        distinct(`GrassPlot ID of plot`)%>% count()
    
    names_datasets_NA <- df %>%
        filter(is.na(`grazed`) & `Grazing (1/0)` == i)%>%
        distinct(`Dataset ID`)%>% pull()
    
    if(nb_plots_0 != 0){
        cat(paste("\nThe dataset(s)", paste(names_datasets_0, collapse = ", "), "contain(s)", 
                    nb_plots_0, "plots with grazed == 0 and Grazing (1/0) =", i))
        cat('\n')
    }
    if(nb_plots_NA != 0){
        cat(paste("\nThe dataset(s)", paste(names_datasets_NA, collapse = ", "), "contain(s)",
                    nb_plots_NA, "plots with grazed = NA and Grazing (1/0) =", i))
        cat('\n')
    }
}
```

#### Manual verification of discrepancies (has to be updated manually if necessary for each version of the masterfile)
##### *Datasets plots with grazed = NA & Grazing (1/0) == '?':*
The following plots in PL_A have grazed = NA and Grazing (1/0) == '?'
```{r echo = F}
df %>%
    filter(is.na(`grazed`) & `Grazing (1/0)` == '?')%>%
    distinct(`GrassPlot ID of plot`)%>% 
    pull()
```
No additional information could be found, so the plots are left as grazed = NA  

##### *Datasets plots with grazed == 0 & Grazing (1/0) == 'probably':*
The following plots in PL_A have grazed == 0 and Grazing (1/0) == ''probably'
```{r echo = F}
df %>%
    filter(`grazed` == 0 & `Grazing (1/0)` == 'probably' )%>%
    distinct(`GrassPlot ID of plot`)%>% 
    pull()
```
No additional information could be found, so the plots are left as grazed == 0.


##### *Datasets plots with grazed == 0 & Grazing (1/0) == '1':*
The following plots in PL_A have grazed == 0 and Grazing (1/0) == 1
```{r echo = F}
df %>%
    filter(`grazed` == '0' & `Grazing (1/0)` == '1' )%>%
    distinct(`GrassPlot ID of plot`)%>% 
    pull()
```
  * PL_A contains 6 plots (PL_A_N005_xxxxb) noted as abandoned with 'occasionally grazed or trampled' in the detailed column (column 96). These plots will be classified as grazed in the new binary column.

```{r}
df[df$`grazed` %in% '0' & df$`Grazing (1/0)` %in% '1',]%<>% 
    mutate(grazed = 1)
```

#### Check
```{r, echo = F}
df%>%
  group_by(`grazed`, `Grazing (1/0)`)%>%
  summarise(n=n())%>%
  spread(`grazed`, n)%>%
  kable(caption = 'Contingency table of columns grazed and Grazing (1/0) after reclassification (check')%>%
  kable_styling(latex_options = "HOLD_position")
```



## Burning
```{r, echo = F}
df%>%
  group_by(`burnt`, `Burning (1/0)`)%>%
  summarise(n=n())%>%
  spread(`burnt`, n)%>%
  kable(caption = 'Contingency table of columns burnt and Burning (1/0)')%>%
  kable_styling(latex_options = "HOLD_position")
```

### Datasets with discrepancies between burnt and Burning (1/0) (automatic report):
```{r, echo = F, results = 'asis'}
unique_binary <- df %>%
    filter(`burnt` == 0 | is.na(burnt))%>%
    distinct(`Burning (1/0)`)%>%
    pull()

unique_binary <- unique_binary[!is.na(unique_binary) & unique_binary != 0]

for(i in unique_binary){
    nb_plots_0 <- df %>%
        filter(`burnt` == 0 & `Burning (1/0)` == i)%>%
        distinct(`GrassPlot ID of plot`)%>% count()
    
    names_datasets_0 <- df %>%
        filter(`burnt` == 0 & `Burning (1/0)` == i)%>%
        distinct(`Dataset ID`)%>% pull()
    
    nb_plots_NA <- df %>%
        filter(is.na(`burnt`) & `Burning (1/0)` == i)%>%
        distinct(`GrassPlot ID of plot`)%>% count()
    
    names_datasets_NA <- df %>%
        filter(is.na(`burnt`) & `Burning (1/0)` == i)%>%
        distinct(`Dataset ID`)%>% pull()
    
    if(nb_plots_0 != 0){
        cat(paste("\nThe dataset(s)", paste(names_datasets_0, collapse = ", "), "contain(s)", 
                    nb_plots_0, "plots with burnt == 0 and Burning (1/0) =", i))
      cat('\n')
    }
    if(nb_plots_NA != 0){
        cat(paste("\nThe dataset(s)", paste(names_datasets_NA, collapse = ", "), "contain(s)",
                    nb_plots_NA, "plots with burnt = NA and Burning (1/0) =", i))
      cat('\n')
    }
}
```

#### Manual verification of discrepancies (has to be updated manually if necessary for each version of the masterfile)
##### *Datasets with burnt == 0 and Burning (1/0) == 1*
The following plots have burnt = 0 and Burning (1/0) == '1':
```{r echo = F}
df %>%
    filter(burnt == 0 & `Burning (1/0)` == '1' )%>%
    distinct(`GrassPlot ID of plot`)%>% 
    pull()
```

No other information are found. These plots are reclassified with burnt == 1

```{r}
df[df$burnt %in% 0 & df$`Burning (1/0)` %in% '1',]%<>%
    mutate(burnt = 1)
```

##### *Datasets with burnt = NA and Burning (1/0) == 1*
The following plots have burnt = NA and Burning (1/0) == '?':
```{r echo = F}
df %>%
    filter(is.na(burnt) & `Burning (1/0)` == '1')%>%
    distinct(`GrassPlot ID of plot`)%>% 
    pull()
```
No other information are found. These plots are reclassified with burnt == 1

```{r}
df[is.na(df$burnt) & df$`Burning (1/0)` %in% '1',]%<>%
    mutate(burnt = 1)
```


##### *Datasets with burnt = NA and Burning (1/0) == ?*
The following plots have burnt == 0 and Burning (1/0) == '?':
```{r echo = F}
df %>%
    filter(burnt ==0 & `Burning (1/0)` == '?' )%>%
    distinct(`GrassPlot ID of plot`)%>% 
    pull()
```
No additional information could be found, so the plots are left as burnt == 0.

#### Check
```{r, echo = F}
df%>%
  group_by(`burnt`, `Burning (1/0)`)%>%
  summarise(n=n())%>%
  spread(`burnt`, n)%>%
  kable(caption = 'Contingency table of columns burnt and Burning (1/0) after reclassification (check)')%>%
  kable_styling(latex_options = "HOLD_position")
```

## Fertilization
Non applicable so far, since the old binary column was an intensity one.

# Proposal for a new land-use template
```{r, echo = F}
# Creer un tableau de recap
```

The column 96 provide a lot of information that is so far not fully exploited.
To make use of it, it requires to manually check each of the value in this column, to fill other columns.
After the Skype meeting (26th), I propose the following columns types:  

## Land-use
Already explained above, these are binary columns providing information on the land-use
There is one extra column that contains verbal information not found in these five columns.  

  * mown
  * grazed
  * burnt
  * fertilized
  * abandoned
  * natural
  * other  

## Land-use intensity
### Mowing
  * mowing_frequency. Numeric, from 0 to infinite. Number of cut(s) per year  
  
### Grazing
  * grazing_intensity. Numeric, from 0 to 1. Self reported relative intensity of grazing.
  * grazing_load. Numeric, from 0 to infinite. Grazing animal load/ha./year (maybe expressed as equivalent cattle)  
  
### Burning
  * burning_frequency. Numeric, from 0 to infinite. Number of burning(s) per year  
  
### Fertilization
  * fertilization_intensity. Numeric, from 0 to 1. Self reported relative intensity of fertilization
  * fertilization_type. Text, 'synthetic' or 'natural'.
  * fertilization_details. Text, free values. Details not present in the tzo previous column (**e.g.** frequency, amount, type of chemical fertilizer...)  


## Abandonement
  * years_abandonment. Numeric, from 0 to infinity. Time since abandonment in years
  * former_land_use: Text, 4 values that can be combined (arable, mown, grazed, burnt). Land-use before abandonment.  

## Grazing details
  * grazing_animal: Text, free values that can be combined, spearated by '/'. New values should be added in agreement with the DB manager.Type of animal grazing the land.   

## Land destination
  * land_destination: Text, 3 possible values (cropland, farmland, recreational). The primary intent behind the land-use.  

## Grassland type
  * primary_grassland: Numeric, 0 or 1. Is the grassland primary (1) or secondary (0)?  


Table: Summary of the proposed template (**to be discussed**)

+------------+--------------------------+-------------+------------------------+
|Column type | Column name              |  Variable   | Possible values        |
|            |                          |type         |                        |
+============+==========================+=============+========================+
| Land use   |- mown                    | - binary    | - 0/1                  |
|            |- grazed                  | - binary    | - 0/1                  |
|            |- burnt                   | - binary    | - 0/1                  |
|            |- fertilized              | - binary    | - 0/1                  |
|            |- abandoned               | - binary    | - 0/1                  |
|            |- natural                 | - binary    | - 0/1                  |
|            |- other                   | - text      | - free                 |
+------------+--------------------------+-------------+------------------------+
| Land-use   |- grazing_intensity       | - numeric   | - 0 to 1               |
| intensity  |- grazing_load            | - numeric   | - 0 to infinity        |
|            |- mowing_frequency        | - numeric   | - 0 to infinity        |
|            |- burning_frequency       | - numeric   | - 0 to infinity        |
|            |- fertilization_intensity | - text      | - 0 to 1               |
|            |- fertilization_type      | - numeric   | - synthetic/natural    |
|            |- fertilization_details   | - text      | - free                 |
+------------+--------------------------+-------------+------------------------+
| Abandonment|- years_abandonment       | - numeric   | - 0 to infinity        |
|            |- former_land_use         | - text      | - arable, mown, grazed |
|            |                          |             | burnt                  |
+------------+--------------------------+-------------+------------------------+
| Grazing    |- grazing animal          | - text      | - free                 |
| details    |                          |             |                        |
+------------+--------------------------+-------------+------------------------+
| Land       |- land_destination        | - text      | - cropland, farmland,  |
|destination |                          |             |recreational            |
+------------+--------------------------+-------------+------------------------+
| Grassland  |- primary_grassland       | - numeric   | - 0/1                  |
|type        |                          |             |                        |
+------------+--------------------------+-------------+------------------------+


## Hamonizing the land-use details column
The following adds the columns not existing yet in the DB, and export the template, with one row per unique value of the land-use detail column.  
This exported file is to be checked manually, in order to fill in the template. To help the sorting, the plots and/or dataset name is left.  
In a second step, it will be reintegrated to the DB, to finalize the harmonization
```{r}
names(df)[which(names(df) == 'Land use detail (e.g. number of cuts or years since abandonment)')] <- "land_use_details"

template_names <- c('mown','grazed','burnt','fertilized','abandoned',
                    'natural','other','grazing_intensity','grazing_load',
                    'mowing_frequency','burning_frequency',
                    'fertilization_intensity','fertilization_type',
                    'fertilization_details','years_abandonment',
                    'former_land_use','grazing animal','land_destination',
                    'primary_grassland')

df_details <- df[match(unique(df$land_use_details), df$land_use_details),]

for(colname in template_names){
  if(!colname %in% names(df_details))
    df_details[,colname] <- NA
}

df_details <- df_details[,c('GrassPlot ID of plot', 'land_use_details', template_names)]


write_xlsx(df_details, path = file.path(path_grassplot, 'df_details.xlsx'))
```


